Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-09-28 article.tex
More updates to Sec 4.2

\paragraph{Case 6: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(0.1, 0.1, 0 \right)$ MPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = 0.04$ m$^2$ and the second moment of area $I = \frac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction was chosen to ensure significant geometric nonlinearity with the upper surface of the beam expected to drop below the plane of the lower surface.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 10 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.

@@

Table \ref{tab:times_memory} lists the wall clock times (time according to a clock on the wall, rounded to the nearest second) and maximum memory usage according to the GNU time utility.
All clock times and memory usage were generated using a Mac Studio with an M2 Ultra CPU on one CPU core, where the code was built with the Clang compiler.
\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
\begin{table}[htb]
	\centering
		\begin{tabular}{ll|ll|ll}
			\hline
			Case & Cell Count & \multicolumn{2}{c|}{JFNK} & \multicolumn{2}{c}{Segregated} \\
			     &            & Time (s) & Memory (MB) & Time (s) & Memory (MB) \\
			\hline 
			MMS & 384 & 0 & $78$ & X & Y \\
				& $4\,374$ & 0 & $86$ & X & Y \\
				& $41\,154$ & 1 & $195$ & X & Y \\
				&  $355\,914$ & 4 & $1\,063$ & X & Y \\
				& $2\,958\,234$ & 41 & $6\,724$ & X & Y \\
				& $24\,118\,074$ & 486 & $30\,606$ & X & Y \\
			\hline
			Elliptic & 45 & 0 & $72$ & X & Y \\
				& 472 & 0 & $78$ & X & Y \\
				& $4\,140$ & 1 & $153$ & X & Y \\
				& $34\,968$ & 7 & $754$ & X & Y \\
				& $287\,280$ & 95 & $3\,943$ & X & Y \\
				& $2\,438\,242$ & 988 & $21\,884$ & X & Y \\
			\hline
			T-member & 624& 0 & 84 & X & Y \\
				& $4\,992$ & 1 & 127 & X & Y \\
				& $39\,936$ & 10 & 557 & X & Y \\
				& $319\,488$ & 154 & $4\,024$ & X & Y \\
				& $2\,555\,904$ & $1\,751$ & $22\,814$ & X & Y \\
			\hline
			Ventricle & $1\,620$ & 54 & 133 & & \\
				& $12\,960$ & 576 & 553 & & \\
				& $103\,680$ & $8\,362$ & $3\,435$ & & \\
				& $829\,440$ & $150\,023$ & $11\,746$ & & \\
			\hline
		\end{tabular}
	\caption{Execution times and maximum memory usage for Jacobian-free Newton-Krylov (JFNK) and segregated methods. Resource requirements are shown for the following cases: manufactured solution (MMS), XX }
	\label{tab:times_memory}
\end{table}

>>>>

\paragraph{Case 6: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(0.1, 0.1, 0 \right)$ MPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = 0.04$ m$^2$ and the second moment of area $I = \frac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction was chosen to ensure significant geometric nonlinearity, with the upper surface of the beam expected to drop below the plane of the lower surface.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 10 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.

@@

Table \ref{tab:times_memory} lists the wall clock times (time according to a clock on the wall, rounded to the nearest second) and maximum memory usage according to the GNU time utility.
All clock times and memory usage were generated using a Mac Studio with an M2 Ultra CPU on one CPU core, where the code was built with the Clang compiler.
\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
\begin{table}[!htb]
	\centering
		\begin{tabular}{ll|ll|ll}
			\hline
			\textbf{Case} & \textbf{Cell Count} & \multicolumn{2}{c|}{\textbf{JFNK}} & \multicolumn{2}{c}{\textbf{Segregated}} \\
			     &            & \textbf{Time} & \textbf{Memory} & \textbf{Time} & \textbf{Memory} \\
			     &            & (in s) & (in MB) & (in s) & (in MB) \\
			\hline 
			\textbf{MMS} & 384 & 0 & $78$ 				& 0 & 79 \\
			\emph{2-D, static,}	& $4\,374$ & 0 & $86$ 			& 0 & 94 \\
			\emph{linear elastic} & $41\,154$ & 1 & $195$  		& 2 & 181 \\
				&  $355\,914$ & 4 & $1\,063$ 		& 9 & $1\,289$ \\
				& $2\,958\,234$ & 41 & $6\,724$ 	& 92 & $6\,174$ \\
				& $24\,118\,074$ & 486 & $30\,606$ & $1\,732$ & $30\,601$ \\
			\hline
			\textbf{Elliptic} & 45 & 0 & $72$ & 0 & 76 \\
			\emph{3-D, static,} & 472 & 0 & $78$ & 0 & 81 \\
			\emph{linear elastic} & $4\,140$ & 1 & $153$ & 8 & 124 \\
				& $34\,968$ & 7 & $754$ & 149 & 660 \\
				& $287\,280$ & 95 & $3\,943$ &  &  \\
				& $2\,438\,242$ & 988 & $21\,884$ &  &  \\
			\hline
			\textbf{T-member} & 624& 0 & 84 					&  &  \\
			\emph{3-D, static,} & $4\,992$ & 1 & 127 				&  &  \\
			\emph{linear elastic} & $39\,936$ & 10 & 557				&  &  \\
				& $319\,488$ & 154 & $4\,024$ 		&  &  \\
				& $2\,555\,904$ & $1\,751$ & $22\,814$ 	&  &  \\
			\hline
			\textbf{Ventricle} & $1\,620$ & 54 & 133 & & \\
			\emph{3-D, static,}& $12\,960$ & 576 & 553 & & \\
			\emph{hyperelastic} & $103\,680$ & $8\,362$ & $3\,435$ & & \\
				& $829\,440$ & $150\,023$ & $11\,746$ & & \\
			\hline
			\textbf{Membrane i} & 9 & 0 & 77 & & \\
			\emph{2-D, static,}& 36 & 0 & 77  & & \\
			\emph{linear elastic}& 144 & 0 & 78  & & \\
				& 576 & 0 & 82 & & \\
				& $2\,304$ & 0 & 94 & & \\
				& $9\,216$ & 1 & 152 & & \\
				& $36\,864$ & 2& 497 & & \\
				& $147\,456$ & 10& $1\,461$ & & \\
			\hline
			\textbf{Membrane ii} & 9 & x & x & & \\
			\emph{2-D, static,} & 36 & x & x & & \\
			\emph{hyperelastic} & 144 & x & x & & \\
				& 576 & x & x & & \\
				& $2\,304$ & x & x & & \\
				& $9\,216$ & x & x & & \\
				& $36\,864$ & x & x & & \\
				& $147\,456$ & & $x$ & & \\
			\hline
			\textbf{Membrane iii} & 9 & x & x & & \\
			\emph{2-D, static,} & 36 & x & x & & \\
			\emph{hyperelastoplastic}& 144 & x & x & & \\
				& 576 & x & x & & \\
				& $2\,304$ & x & x & & \\
				& $9\,216$ & x & x & & \\
				& $36\,864$ & x & x & & \\
				& $147\,456$ & & $x$ & & \\
			\hline
			\textbf{Cantilever} & $270$ & 11 & 85 & & \\
			\emph{3-D, dynamic,} & $2\,160$ & 71 & 181 & & \\
			\emph{hyperelastic}& $17\,280$ & 854 & 752 & & \\
				& $138\,240$ & $11\,800$ & $5\,790$ & & \\
			\hline
			\textbf{Turbine} & $x$ & x & x & & \\
			\emph{3-D, static,} & & & & & \\
			\emph{linear elastic} & & & & & \\
			\hline
			\textbf{FSI} & $1\,252$ + 156 & $4\,265$ & 743 & $4\,820$ & 752 \\
			\emph{2-D, dynamic,}	& $5\,008$ + 624 & $33\,366$ & 788 & $46\,328$ & 786 \\
			\emph{hyperelastic} & $20\,032$ + $2\,496$ & $154\,237$ & 729 & & \\
			\hline
		\end{tabular}
	\caption{Execution times and maximum memory usage for Jacobian-free Newton-Krylov (JFNK) and segregated methods. Resource requirements are shown for the following cases: manufactured solution (MMS), XX }
	\label{tab:times_memory}
\end{table}
